---
title: "Validating Community Assembly Rules Project Data"
output: rmarkdown
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# CollaboratoR: Example Validation Session

This is an example workflow for using the CollaboratoR package for validation of data files from the project for which is was originally created.  To run this requires google drive credentials in the .Renviron file or the 
environment.   

This notebook is just a demonstration and has some specific configuration for that project, such as multi-tab data sheets in google drive.   

For those with limited or low-tier Google workspace accounts, there is a rate limit on reading googlesheets that
may require several minutes for this R-markdown file to render. 

```{r setup}
require(dplyr)
library(CollaboratoR)
```

## Authenticate to gdrive/sheets.  Requires an email account

```{r gdrive setup}

## EDIT AND ADD YOUR EMAIL IF YOU DON'T HAVE AN EMAIL IN .Renviron FILE
# if you have not setup the .Renvfile, or would like to test a different email 
# address, set it here.   If none is sent, the auth setup function will look in 
# the environment (e.g. .Renviron file)
drive_email <- Sys.getenv('PROJECT_EMAIL')
authenticated <- CollaboratoR::gsheet_auth_setup(drive_email = drive_email)
if(!authenticated) {
  print("EDIT AND ADD YOUR EMAIL IF YOU DON'T HAVE AN EMAIL IN .Renviron FILE")
  stop()
} 
print(paste("Authenticaed with ", drive_email))
```



## Load data specific to the Comm Assembly Rules project stored in this package

```{r, start, echo=TRUE}
# Validate package rules
data(commassembly_rules_biomass_str)
data(commassembly_rules_env_str)

# list of URLS from google drive to a data sheet
doc_with_list_url <- Sys.getenv('TEST_ID_LIST_URL')
id_column = 'ID_new'
urls.df <- read_url_list(gurl = doc_with_list_url, id_column = id_column, url_column='url')
print(paste(nrow(urls.df), "urls to read"))
```

Read in the first URL as a test.  There will be warnings if the columns are not the correct type or name
```{r set_example_url, echo=TRUE}
# get the example from one of the URLS

test_gsheet_url <- Sys.getenv('TEST_URL')
if(test_gsheet_url == "") {
  sheet_count <- nrow(urls.df)

  test_url_num <- sample(1:sheet_count, 1)
  test_gsheet_url <- sample(urls.df$url, test_url_num)
}

print(paste("We'll use this test sheet:", test_gsheet_url))

```

#### Read in data from shee

This project's data sheets have two tabs in each: one for biomass and one for Env

Read in biomass tab into data frame
```{r, read_one_biomass, echo=TRUE}
# test of basic reading, and validate the column names
# this will throw warnings if there are invalid column names

biomass.df <- read_data_sheet(gurl = test_gsheet_url, 
                              tab_name = 'biomass_data', 
                              spec.df = commassembly_rules_biomass_str)

biomass.df
```

Read in "env" tab

```{r, read_one_env, echo=TRUE}
env.df <- read_data_sheet(gurl = test_gsheet_url, 
                          tab_name = 'env_data', 
                          spec.df = commassembly_rules_env_str)

if(typeof(env.df) == 'list' && nrow(env.df) > 0) {
  print(paste("successfully read env.df from ", test_gsheet_url, "tab = env_data"))
  head(env.df)
} else {
  print(paste("problem reading env.df from ", test_gsheet_url, "tab = env_data"))
}


```

## Validation

the above does validate the column headers and types.  (TDB problems output from readr)

Let's try using the validation rules

```{r, validate biomass from file, echo=TRUE, error=TRUE}

biomass_validation_file <-  here::here('inst/rules/biomass_validation_rules.yaml')
if ( file.exists(biomass_validation_file)) {
  validation_object <- validate_from_file(biomass.df, biomass_validation_file)
  validation_summary <- validate::summary(validation_object)
  print(validation_summary)

} else {
  print("could not find validation rules file")
}

```

Validation: Just show the fails, if any
```{r, show validation fails}
if(sum(validation_summary$fails) == 0) {
  print("Validation Passed, return true")
} else {
  fails <- validation_summary[validation_summary$fails > 0,]
  fails
}

```

**Validate the 'env' tab**

```{r, validate env from file, echo=TRUE, error=TRUE}

read_data_sheet_save_warnings <- errorSaver(read_data_sheet)
env.df <- read_data_sheet_save_warnings(gurl = test_gsheet_url, 
                              tab_name = 'env_data', 
                              spec.df = commassembly_rules_env_str)

if("warnings" %in% names(env.df)) { 
  print(env.df$warnings)
  # get just the data
  env.df <- env.df[[1]]
}

```

```{r}
env_validation_file <- here::here('inst/rules/env_validation_rules.yaml')

if(file.exists(env_validation_file)) {
  validation_object <- validate_from_file(env.df, env_validation_file)
  validation_summary <- validate::summary(validation_object)
  if(sum(validation_summary$fails) == 0) {
    print("Validation Passed, return true")
  } else {
    print("Validation fails")
    fails <- validation_summary[validation_summary$fails > 0,]
    print(fails)
  }
} else {
  print(paste("could not find validation rules file", env_validation_file))
}

```


## Function to read, check columns, validate and save the CSVS of these two tabs

```{r, functions to write sheets, echo=FALSE, error=TRUE}
csv_folder = '../L0'
dir.create(csv_folder, showWarnings = FALSE)
biomass_validation_file <- '../inst/rules/biomass_validation_rules.yaml'
env_validation_file <- '../inst/rules/env_validation_rules.yaml'

# sheet_data is one row of the URLS data frame, as a list, url is ..$url
read_and_report<- function(url, tab_name, spec.df){
  read_data_sheet_save_warnings <- CollaboratoR::errorSaver(read_data_sheet)
  data.df <- read_data_sheet_save_warnings(gurl = url, 
                              tab_name, 
                              spec.df)

  if("warnings" %in% names(data.df)) { 
      print(paste("reading warnings for tab_name:", tab_name, "url", url))
      print(data.df$warnings)
      # get just the data so what's left can be validated
      data.df <- data.df[[1]]
  }
    
  return(data.df)
}


save_csvs<- function(sheet_data){
  # print(paste(sheet_data$id, sheet_data$url))

  url <- sheet_data$url
  
  #### ENV  
  env.df <- read_and_report(url, 
                            tab_name = 'env_data', 
                            spec.df = commassembly_rules_env_str)
  if(!exists('env.df')) {
    print(paste("could not read sheet: "))
    return (c(NA, NA))
  } 
  env_valid <- validation_report(env.df, env_validation_file)

  ## BIOMASS READ
  biomass.df <- read_and_report(url, 
                                tab_name = 'biomass_data', 
                                spec.df = commassembly_rules_biomass_str)
  if(!exists('biomass.df')) {
    print(paste("could not read sheet "))
    return (c(NA, NA))
  } 
  # BIOMASS VALIDATION
  biomass_valid <- validation_report(biomass.df, biomass_validation_file)
   
  if(biomass_valid && env_valid) {
    # SAVE
    id_new <- sheet_data$ID_new
    biomass_file_name <- file.path(csv_folder, paste0('biomass_', id_new, '.csv'))
    utils::write.csv(biomass.df, biomass_file_name, row.names = FALSE)
    env_file_name <- file.path(csv_folder, paste0('env_', id_new, '.csv'))
    utils::write.csv(env.df, env_file_name, row.names = FALSE)
    return(c(biomass_file_name, env_file_name))
  } else {
    print(paste("did not validate biomass", biomass_valid, " env ", env_valid ))
    return(c(NA,NA))
  }
}


```

###  Save one to a csv and output the filename

```{r, save_one, echo=TRUE, error=TRUE}

test_url_num <- sample(1:nrow(urls.df), 1)
test_gsheet <- as.list(urls.df[test_url_num,])
print(test_gsheet)
test_file_names<- save_csvs(test_gsheet)

print(test_file_names)
```

### Test of our prcoess

Try to read the CSV back in and show it


```{r}
test_df <- read_data_csv(test_file_names[1], spec.df = commassembly_rules_biomass_str)
head(test_df)
```

if that worked, move on to the next step:

## Loop that reads all URLS in the url list. 

This is complete code, with duplicates from above, to show how to validate all CSVs

```{r, save them all, echo=TRUE, error=TRUE}

data(commassembly_rules_biomass_str)
data(commassembly_rules_env_str)

# before running this, authenticate. 
# if you don't have ths PROJECT_EMAIL set in .Renviron, edit the next line
drive_email = Sys.getenv('PROJECT_EMAIL')
stopifnot(gsheet_auth_setup(drive_email = drive_email))


# list of URLS from google drive to a data sheet
doc_with_list_url <- Sys.getenv('TEST_ID_LIST_URL')
id_column = 'ID_new'
urls.df <- read_url_list(gurl = doc_with_list_url, id_column = id_column, url_column='url')

for(i in seq(1:nrow(urls.df))) {
  print(paste(urls.df$ID_new[i], urls.df$id[i], urls.df$who[i]))
  tryCatch({
    study_info <- as.list(urls.df[i,])
    file_names<- save_csvs(study_info)
    print(file_names)
    }, error=function(e) print(e))
}
```



## Post-Validation example: build a database

This reads just the saved CSV from the 'biomass' tab and aggregates
into a CSV.  This methos requires that all of the CSVs have the same 
columns, meaning all of the Google sheets read above have the same columns, 
which they should since the've passed validation. 

```{r, combine, echo=TRUE}
# getting the list of CSV file paths is flexible so that some could be 
# excluded or multiple folders combined

biomass_files <- dir('../L0', pattern = "biomass.*\\.csv", 
                          full.names = TRUE, include.dirs = TRUE)
biomass.df <- aggregate_csvs(csv_list=biomass_files, spec.df=commassembly_rules_biomass_str)

print(paste('biomass rows', nrow(biomass.df))) # , " env rows", nrow(env.df)))
```
