---
title: "Project Configuration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Project Configuration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(CollaboratoR)
```

As described in the Vignette describing the [Google Workspace API](google_sheets_api.html), 
this package needs to find a handful of api keys and credentials files.   In addition, this pacakge
will look for some of the base files for it's workflow.  

To save those values to be re-used quickly, rather than put them into your code directly,  R
has a way to inject those values in to the OS environment, using the `.Renviron` file. 

See [Dr. Jenny Bryan's chapter on this file and R start system](https://rstats.wtf/r-startup.html#renviron) 
for a description of this file. 

Here is an example of values to go into the file.  Copy the section below into a new file 
and save as `.Renviron`  

Do not put any spaces on either side of the `=` when assigning values.  For example use
`x="x"` syntax.  This is not an **R** script file, but an operating system file read by the R 
system when it starts up.  

```
# .Renviron for project configuration. 

# save this example file with the name .Renviron 
### Google Cloud Workspace API settings for private workspaces
PROJECT_EMAIL='your email here'  # email used to log-in to google sheets
PROJECT_API_KEY='AIzaSyDzkJvGiW6U0RsjWh9fFqxxlucFCJIeOyc'
DRIVE_API_ID_NAME='my_api_resource'

### Google Workspace Drive names and locations

# name of the shared drive create in Google Workspace that team members have access to
DRIVE_API_ID_FILE='path/to/gcp/credentials.json'

#name of API id created inside GCP
PROJECT_SHARE_DRIVE='mysharedrive'

# path inside the shared drive for main L0 data sheets, fo
PROJECT_SHARE_DRIVE_PATH=''

### URIs for Google Sheets or CSV files for testing (URLs will look like examples)
TEST_URL='https://docs.google.com/spreadsheets/d/uniqueIDhere'
TEST_ID_LIST_URL='https://docs.google.com/spreadsheets/d/sheet-id-value-here'
TEST_TEMPLATE_URL='https://docs.google.com/spreadsheets/d/sheet-id-value-here'
```